Exploring the Design Space of Differentially Private Bayesian Inference under Hellinger Distance

### 2019-01-23 ICML2019/main.tex
working on section 3 and 4

\noindent \textbf{Main contributions.}
What follows is a summary of the main contributions of this work.
\begin{itemize}
	\item We design a new differentially private Bayesian inference mechanism based on the standard exponential mechanism.
    \item A full proof of $(\epsilon, \delta)$-differential privacy of the mechanism is given in the paper. \gp{it's only eps-d.p for now}
	\item We explored two ways to improve the accuracy: 1) calibrating noise to the sensitivity of a metric over distributions
      (e.g. Hellinger distance ($\hellinger$), $f$-divergences, etc\dots). 2) Using a smooth upper bound on the local sensitivity and scale
      the noise to this smooth bound rather than global sensitivity, to improve the mechanism accuracy.
  \item We implemented the new proposed mechanism and other art-of-state mechanisms, comparing their performances in terms of accuracy and privacy.
\end{itemize}

@@

@@

\noindent \emph{A note on input data representation.}
We will think of our input databases $\dataobs$ of length $n$ as histograms 
with $k$-$1$ bins where $k$ is the dimensionality of the problem.
This is w.l.o.g. because given the first $k$-$1$ counts, the $k$-th count
is known.

\section{Problem Statement}
\label{subsec_problem}
We are interested in exploring the design space of mechanisms for privately releasing the posterior
distribution for the Dirichlet-Multinomial model as described in the
previous section.
It is worth noticing that the posterior distribution of this model is fully characterized
by its parameters. That is, we are interested in releasing a private version of
the Dirichlet distribution with parameter
$\valpha'=(\valpha + \Delta \valpha)$, or in the case of the
Beta-Binomial model we are interested in releasing a private version of the Beta distribution with 
parameters $(\alpha',\beta')=(\alpha + \Delta \alpha,\beta + n -
\Delta \alpha)$.
\citet{zhang2016differential} and \citet{xiao2012bayesian}
have already attacked this problem by adding independent Laplacian noise to the
parameters of the posteriors. That is, in the case of the Beta-Binomial system,
the values released would be: $(\tilde\alpha,\tilde\beta)=(\alpha +  \widetilde{\Delta \alpha},\beta + n - \widetilde{\Delta \alpha})$
where $\widetilde{\Delta \alpha}\thicksim \lap{\Delta \alpha}{\frac{2}{\epsilon}}$,
and where $\lap{\mu}{\nu}$ denotes a Laplace random variable with mean $\mu$ and scale $\nu$.
This mechanism is $\epsilon$-differentially private, and the noise is
calibrated w.r.t. to a sensitivity of 2 which is derived by using
$\ell_1$ norm over the pair of parameters. Indeed, considering two
adjacent data observations $\dataobs, \dataobs'$, that, from a unique prior, give rise to two posterior
distributions, characterized by the pairs
$(\alpha',\beta')$ and $(\alpha'',\beta'')$ then
$|\alpha'-\alpha''|+|\beta'-\beta''|\leq 2$.
This argument extends similarly to the Dirichelet-Multinomial system. Details are introduced in Section \ref{sec_mechs}.\\
However, in previous works, the accuracy of the posterior was measured with respect to $\ell_1$ norm. That is, an upper bound was given on
\[
  \Pr[|\alpha - \tilde\alpha| + |\beta - \tilde\beta |\geq \gamma ]
\]
where $(\alpha, \beta), (\tilde\alpha,\tilde\beta)$ are the real and the perturbed pairs of parameters, respectively.
However, distance between
distributions are more naturally  measured by distribution
metrics such as $f$-divergences, e.g. Hellinger distance, etc., than the $\ell_1$ norm shown above.
{\color{blue} still not sure this is a good way to put it...}
This gives us motivation for exploring the design space
of mechanisms by considering different norms. e.g. a norm over space of measures.
to compute the sensitivity and provide guarantees on the accuracy.

We will use the \emph{Hellinger} distance $\hellinger(\cdot,\cdot)$.
Given two probability distributions $\textup{P}$ and $\textup{Q}$,
the square of the Hellinger distance between them is defined as:
\[
\hellinger(P,Q)^2 =\frac{1}{2} \int \bigg(\frac{dP}{d \lambda} - \frac{dQ}{d \lambda}\bigg)^2 d \lambda
\]
When $\textup{P}=\dirichlet{\valpha_1}$ and $\textup{Q}=\dirichlet{\valpha_2}$, we can unfold the above and derive: 
 \[
   \hellinger(\dirichlet{\valpha_1}, \dirichlet{\valpha_2}) = \sqrt{1 - \frac{\betaf(\frac{\valpha_1 + \valpha_2}{2})}{\sqrt{\betaf(\valpha_1) \betaf(\valpha_2)}}}
 \]
Our choice to use Hellinger distance is motivated by three facts:
\begin{itemize}
  \item It simplifies calculations in the case of the probabilistic models considered here.

@@

\section{Private Mechanisms}
\label{sec_mechs}
In this section we will describe and analyse a series of mechanisms.
Section \ref{sec_ell1_dim} presents a first mechanism, $\lapmech$,
based on the Laplace primitive where the sensitivity is measured w.r.t. to $\ell_1$ norm.
Section \ref{sec_ell1_hist}, shows that $\lapmech$'s accuracy can be improved
by performing a better analysis on the sensitivity of the inference process.
Section \ref{sec_ehd}, shows an instance of the exponential mechanism, $\expmech$, where the scoring
function is based on the Hellinger distance, we will see how this mechanisms performs
poorly due to the very high global sensitivity of the inference process w.r.t to this metric.
Section \ref{sec_ehdl} introduces, $\lexpmech$, a variation of $\expmech$ which uses a lower bound
on the global sensitivity called local sensitivity.
Unfortunately, despite improving on accuracy $\lexpmech$ is non-private.
Finally, building on $\lexpmech$, Section \ref{sec_ehds} introduces $\hexpmech$:
which, based on the notion of \emph{smooth} sensitivity, achieves $\epsilon$-differential privacy and
has good accuracy in specific cases.
In presenting these mechanisms we will denote by $\inferd{\valpha}{\vtheta}{\dataobs}$ the posterior  $\dirichlet{\valpha'}$
obtained starting from the prior $\dirichlet{\valpha}$, likelihood $\multinomial{|\dataobs|}{\vtheta}$, and observed data $\dataobs$.
To minimize notation we will identify the distribution $\dirichlet{\valpha}$ with its vector
of parameters.
Two different upper bounds to the value of
\begin{align}
  \label{eqn:eq1}
  \max_{\substack{\adj{\dataobs}{\dataobs'},\\
      \valpha=\inferd{\valpha}{\vtheta}{\dataobs},\\
      \valpha'=\inferd{\valpha}{\vtheta}{\dataobs'}}}||\valpha-\valpha'||_1 
\end{align}
will be used to calibrate the noise in the mechanisms based on the Laplacian primitive.
Finally, with $\bysinfer{\dataobs}$, and $\bys{\dataobs}{\vtheta}$ we will denote $\inferd{\valpha}{\vtheta}{\dataobs}$ when $\valpha, \vtheta$ or just $\valpha$
are known from the context.

\subsection{Calibrating noise w.r.t. $\ell_1$ norm ($\lapmech$)}
\label{sec_ell1_dim}
We start off by presenting $\lapmech$, in Algorithm \ref{lapmech}, based on the Laplace primitive.
The idea of $\lapmech$ is to add Laplacian noise with scale proportional to $\mathcal{X}$ to
the differences between the respective numeric parameters of the prior
and posterior distribution. 
\begin{algorithm}
  \caption{$\lapmech$}
  \label{lapmech}
  \begin{algorithmic}
  \INPUT $\dataobs\in\mathcal{X}^n$, $\dirichlet{\valpha}$
  \STATE \quad {\bf let} $\valpha' = \inferd{\valpha}{\vtheta}{\dataobs}$
  \STATE \quad {\bf Initialize} a vector $\tilde{\valpha} = (0, \dots, 0 )\in \mathbb{N}^{|\mathcal{X}|}$ 
  \STATE \quad {\bf For} $i = 1 \dots |\mathcal{X}| - 1$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{|\mathcal{X}|}{\epsilon}}$
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_{|\mathcal{X}|}} = \alpha_{|\mathcal{X}|} + \lfloor n - \sum_{i = 1}^{|\mathcal{X}|-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \STATE {\bf return} $\tilde{\valpha}$
  \end{algorithmic}
\end{algorithm}
\begin{lem}
  Algorithm \ref{lapmech} is $\epsilon$-d.p.
  \begin{proof} By noticing that $\mathcal{X}-1$ is an upper bound to the value denoted by (\ref{eqn:eq1}). \end{proof}
\end{lem}

\subsection{Improving the sensitivity ($\ilapmech$)}
\label{sec_ell1_hist}
By seeing the parameters of the distributions as histograms, we can give a tighter upper bound to the value denoted by (\ref{eqn:eq1}).
Indeed a constant will do. Algorithm $\ilapmech$ in Algorithm \ref{ilapmech} takes advantage of this fact. 
Indeed, when $\mathcal{X}=2$, 1 is an upper bound for the value of (\ref{eqn:eq1}), while when $\mathcal{X}>2$, 2 is.
In Algorithm \ref{ilapmech} the algorithm $\ilapmech$ is presented.
\begin{algorithm}
  \caption{$\ilapmech$}
  \label{ilapmech}
  \begin{algorithmic}
  \INPUT $\dataobs\in\mathcal{X}^n$, $\dirichlet{\valpha}$
  \STATE \quad {\bf let} $\valpha' = \inferd{\valpha}{\vtheta}{\dataobs}$
  \STATE \quad ${\bf let}\ k=\left \{
        \begin{array}{lll}          
          1&\text{if}& |\mathcal{X}|=2\\
          2&\text{otherwise}
        \end{array}
      \right .\ $
  \STATE \quad {\bf Initialize} a vector $\tilde{\valpha} = (0, \dots, 0 )\in \mathbb{N}^{|\mathcal{X}|}$ 
  \STATE \quad {\bf For} $i = 1 \dots |\mathcal{X}|-1$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{k}{\epsilon}}$
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_{|\mathcal{X}|}} = \alpha_{|\mathcal{X}|} + \lfloor n - \sum_{i = 1}^{|\mathcal{X}|-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \STATE {\bf return} $\tilde{\valpha}$
  \end{algorithmic}
\end{algorithm}
\begin{lem}
  Algorithm \ref{ilapmech} is $\epsilon$-d.p.
  \begin{proof} By noticing that 1 and 2 are upper bounds to the value denoted by (\ref{eqn:eq1}), respectively when $|\mathcal{X}|=2$ and $|\mathcal{X}|>2$.
  \end{proof}
\end{lem}

\subsection{A new family of mechanisms}
\label{sec_exp}
In this section, we explore  mechanism which we denote by $\expmech$. $\expmech$ is an instance
of the exponential mechanism \cite{talwar} where the scoring function is based on a metric over
probability distributions.
The idea is that $\expmech$ will sample and output a distribution with higher probability when
it is closer to the real posterior; importantly,  \emph{closer} is measured using a metric over probability
distributions as opposed to a metric over the parameters that represent it. In this work we will
use the Hellinger distance, but this approach could be extended to other $f$-divergences. 
Given a prior distribution $\dirichlet{\valpha}$,
we define the follwing set  where the mechanism will sample from:
\[
  \candidateset=\{\inferd{\valpha}{\vtheta}{\dataobs}\mid \dataobs \in \mathcal{X}^n\}.
\]
In general $|\candidateset|=\binom{n + 1}{|\mathcal{X}| - 1}$, and in the specific case of $|\mathcal{X}|=2$, that is the Beta-Binomial model,
the above reduces to:
\[
  \{(\alpha',\beta')\mid \alpha'=\alpha + \Delta\alpha, \beta'=\beta + n-\Delta\alpha,  \Delta\alpha=\sum x_i, \dataobs\in \{0,1\}^n\}
\]
where $\valpha=(\alpha,\beta)$.

\noindent \textbf{A first approach: the $\expmech$ mechanism.} $\expmech$'s pseudocode is shown in Algorithm \ref{mech:expmech1}.
Given the input $\dataobs\in\{0,1\}^n$ and a prior $\betad(\alpha, \beta)$, the mechanism $\expmech$ outputs an element $r\in\candidateset$ with  probability: 
\[
\frac
  {\exp(\frac{\epsilon\cdot \ux{r}}{2\cdot GS})}
{\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot \ux{r'}}{2\cdot GS})},
\]
where $u(\dataobs,r)$ is the Hellinger based score function:
$-\hellinger(\bysinfer{\dataobs}, r)$, and $GS$ is the global sensitivity of the scoring function, defined as:
\[
GS = 
\max_{\dataobs,\dataobs': \\ \adj{\dataobs}{\dataobs'}}\max_{\{r\in \mathcal{R}\}}
|\hellinger(\bysinfer{\dataobs}, r) - \hellinger(\bysinfer{\dataobs'}, r)|
\]
% 
  \begin{algorithm}
  \caption{$\expmech$}
  \label{mech:expmech1}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \mathcal{X}^n$, prior: $\dirichlet{\valpha}$, $\epsilon$
  \STATE \quad {\bf let} $\dirichlet{\valpha'}= \bys{\dataobs}{\valpha}$.   
  \STATE \quad {\bf Initialize} associative map $pr$ of size $|\candidateset|$.
  \STATE \quad {\bf For} $r$ in $\candidateset$:
  \STATE \quad \quad $pr[r] = 
  {\exp(\frac{\epsilon\cdot \ux{r}}{2\cdot GS})}
/ {\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot u(\dataobs, r')}{2\cdot GS})}$
\STATE \quad {\bf sample} in $z$ from probability distribution $pr$.
\STATE {\bf return} $z$
  \end{algorithmic}
  \end{algorithm}
\begin{lem}
  Algorithm \ref{mech:expmech1} is $\epsilon$-d.p.
\end{lem}

\noindent \textbf{A second approach based on local sensitivity: $\lexpmech$.} Figure \ref{fig_theory} shows that  the accuracy of
$\expmech$ is poor if compared with the one of $\ilapmech$. This is due to the fact that
$GS$ is relatively large and hence the variance of the distribution $\expmech$ ends up also being too large. In order to solve this issue we
will proceed in two steps. We will first define, in Algorithm   \ref{mech:lexpmech}, $\lexpmech$, which instead of using the global sensitivity, will use the local sensitivity to calibrate
the noise. This is known to be a non private approach, but it will guide us towards new insights.
We will then proceed to use a smooth bound the local sensitivity to
calibrate the noise.

  \begin{algorithm}
  \caption{$\lexpmech$}
  \label{mech:lexpmech}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \mathcal{X}^n$, prior: $\dirichlet{\valpha}$, $\epsilon$
  \STATE \quad {\bf let} $\dirichlet{\valpha'}= \bys{\dataobs}{\valpha}$.   
  \STATE \quad {\bf Initialize} associative map $pr$ of size $|\candidateset|$.
  \STATE \quad {\bf For} $r$ in $\candidateset$:
  \STATE \quad \quad $pr[r] = 
  {\exp(\frac{\epsilon\cdot \ux{r}}{2\cdot LS(\dataobs)})}
/ {\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot u(\dataobs, r')}{2\cdot LS(\dataobs)})}$
\STATE \quad {\bf sample} in $z$ from probability distribution $pr$.
\STATE {\bf return} $z$
  \end{algorithmic}
  \end{algorithm}

>>>>

\noindent \textbf{Main contributions.}
What follows is a summary of the main contributions of this work.
\begin{itemize}
	\item We design a new differentially private Bayesian inference mechanism based on the standard exponential mechanism.
    \item A full proof of $(\epsilon, \delta)$-differential privacy of the mechanism is given in the paper. 
	\item We explored two ways to improve the accuracy: 1) calibrating noise to the sensitivity of a metric over distributions
      (e.g. Hellinger distance ($\hellinger$), $f$-divergences, etc\dots). 2) Using a smooth upper bound on the local sensitivity and scale
      the noise to this smooth bound rather than global sensitivity, to improve the mechanism accuracy.
  \item We implemented the new proposed mechanism and other art-of-state mechanisms, comparing their performances in terms of accuracy and privacy.
\end{itemize}

@@

\noindent \textbf{Hellinger distance.}
\emph{Hellinger} distance $\hellinger(\cdot,\cdot)$ is a statistical
divergence measuring the distance between two probability
distributions. Given two probability distributions $\textup{P}$ and $\textup{Q}$,
the Hellinger distance between them is defined as:
\[
\hellinger(P,Q) =\sqrt{\frac{1}{2} \int \bigg(\frac{dP}{d \lambda} - \frac{dQ}{d \lambda}\bigg)^2 d \lambda}
\]
When $\textup{P}=\betad(\alpha_1, \beta_1)$ and $\textup{Q}=\betad(\alpha_2, \beta_2)$, we can unfold the above and derive: 
\[
 \hellinger(P,Q) =  \sqrt{1 - \frac{\betaf(\frac{\alpha_1 + \alpha_2}{2}, \frac{\beta_1 + \beta_2}{2})}{\sqrt{\betaf(\alpha_1,\beta_1)\betaf(\alpha_2,\beta_2)}}}
\]

@@

\noindent \emph{A note on input data representation.}
We will represent our input databases $\dataobs$ of length $n$ as histograms 
with $k$-$1$ bins where $k$ is the dimensionality of the problem.
This is w.l.o.g. because given the first $k$-$1$ counts, the $k$-th count
is known.

\section{Problem Statement}
\label{subsec_problem}
We are interested in exploring the design space of mechanisms for privately releasing the posterior
distribution for the Dirichlet-Multinomial model by output
perturbation with accuracy measured in Hellinger distance.
It is worth noticing that the posterior distribution of this model is fully characterized
by its parameters. That is, we are interested in releasing a private version of
the Dirichlet distribution with parameter
$\valpha'=(\valpha + \Delta \valpha)$, where $\valpha$ is the
parameter of the prior, and $\Delta \valpha$ is computed from the data
as discussed in the previous section. In the case of the
Beta-Binomial model this specialize to releasing a private version of the Beta distribution with 
parameters $(\alpha',\beta')=(\alpha + \Delta \alpha,\beta + n -
\Delta \alpha)$, where $\alpha$ and $\beta$ are the parameters of the
prior, and $\Delta \alpha$ is the number of $1$ in the data.

Previous work by \citet{zhang2016differential} and \citet{xiao2012bayesian}
have considered the problem of releasing the posterior of the
Beta-Binomial model by output perturbation with accuracy measured in terms of
$\ell_1$ norm. In addition,   \citet{zhang2016differential} has also
considered KL-divergence between posteriors. 
In this work we consider instead the Hellinger Distance between
posteriors. Similarly to KL-divergence, Hellinger Distance is an
$f$-divergence measuring levels of similarities between
distributions. However, unlikely KL-divergence, Hellinger Distance is a proper
distance satisfying reflexivity, symmetry and triangle
inequality. These properties gives Hellinger Distance analytical
advantages and makes the minimum hellinger distance  an efficient and
robust estimator~\cite{}.

attacked this problem by adding independent Laplacian noise to the
parameters of the posteriors. That is, in the case of the Beta-Binomial system,
the values released would be: $(\tilde\alpha,\tilde\beta)=(\alpha +  \widetilde{\Delta \alpha},\beta + n - \widetilde{\Delta \alpha})$
where $\widetilde{\Delta \alpha}\thicksim \lap{\Delta \alpha}{\frac{2}{\epsilon}}$,
and where $\lap{\mu}{\nu}$ denotes a Laplace random variable with mean $\mu$ and scale $\nu$.
This mechanism is $\epsilon$-differentially private, and the noise is
calibrated w.r.t. to a sensitivity of 2 which is derived by using
$\ell_1$ norm over the pair of parameters. Indeed, considering two
adjacent data observations $\dataobs, \dataobs'$, that, from a unique prior, give rise to two posterior
distributions, characterized by the pairs
$(\alpha',\beta')$ and $(\alpha'',\beta'')$ then
$|\alpha'-\alpha''|+|\beta'-\beta''|\leq 2$.
This argument extends similarly to the Dirichelet-Multinomial system. Details are introduced in Section \ref{sec_mechs}.\\
However, in previous works, the accuracy of the posterior was measured with respect to $\ell_1$ norm. That is, an upper bound was given on
\[
  \Pr[|\alpha - \tilde\alpha| + |\beta - \tilde\beta |\geq \gamma ]
\]
where $(\alpha, \beta), (\tilde\alpha,\tilde\beta)$ are the real and the perturbed pairs of parameters, respectively.
However, distance between
distributions are more naturally  measured by distribution
metrics such as $f$-divergences, e.g. Hellinger distance, etc., than the $\ell_1$ norm shown above.
{\color{blue} still not sure this is a good way to put it...}
This gives us motivation for exploring the design space
of mechanisms by considering different norms. e.g. a norm over space of measures.
to compute the sensitivity and provide guarantees on the accuracy.

Our choice to use Hellinger distance is motivated by three facts:
\begin{itemize}
  \item It simplifies calculations in the case of the probabilistic models considered here.

@@

\section{Private Mechanisms}
\label{sec_mechs}
In this section we will describe and analyse a series of mechanisms.
Section \ref{sec_ell1_dim} presents a first mechanism, $\lapmech$,
based on the Laplace primitive where the sensitivity is measured w.r.t. to $\ell_1$ norm.
Section \ref{sec_ell1_hist}, shows that $\lapmech$'s accuracy can be improved
by performing a better analysis on the sensitivity of the inference process.
Section \ref{sec_ehd}, shows an instance of the exponential mechanism, $\expmech$, where the scoring
function is based on the Hellinger distance, we will see how this mechanisms performs
poorly due to the very high global sensitivity of the inference process w.r.t to this metric.
Section \ref{sec_ehdl} introduces, $\lexpmech$, a variation of $\expmech$ which uses a lower bound
on the global sensitivity called local sensitivity.
Unfortunately, despite improving on accuracy $\lexpmech$ is non-private.
Finally, building on $\lexpmech$, Section \ref{sec_ehds} introduces $\hexpmech$:
which, based on the notion of \emph{smooth} sensitivity, achieves $\epsilon$-differential privacy and
has good accuracy in specific cases.
In presenting these mechanisms we will denote by $\inferd{\valpha}{\vtheta}{\dataobs}$ the posterior  $\dirichlet{\valpha'}$
obtained starting from the prior $\dirichlet{\valpha}$, likelihood $\multinomial{|\dataobs|}{\vtheta}$, and observed data $\dataobs$.
Two different upper bounds to the value of
\begin{align}
  \label{eqn:eq1}
  \max_{\substack{\adj{\dataobs}{\dataobs'},\\
      \dirichlet{\valpha}=\inferd{\valpha}{\vtheta}{\dataobs},\\
      \dirichlet{\valpha'}=\inferd{\valpha}{\vtheta}{\dataobs'}}}||\valpha-\valpha'||_1 
\end{align}
will be used to calibrate the noise in the mechanisms based on the Laplacian primitive.

\subsection{Calibrating noise w.r.t. $\ell_1$ norm ($\lapmech$)}
\label{sec_ell1_dim}
We start off by presenting $\lapmech$, in Algorithm \ref{lapmech}, based on the Laplace primitive.
The idea of $\lapmech$ is to add Laplacian noise with scale proportional to $\mathcal{X}$ to
the differences between the respective numeric parameters of the prior
and posterior distribution. 
\begin{algorithm}
  \caption{$\lapmech$}
  \label{lapmech}
  \begin{algorithmic}
  \INPUT $\dataobs\in\mathcal{X}^n$, $\dirichlet{\valpha}$
  \STATE \quad {\bf let} $\dirichlet{\valpha'} = \inferd{\valpha}{\vtheta}{\dataobs}$
  \STATE \quad {\bf Initialize} a vector $\tilde{\valpha} = (0, \dots, 0 )\in \mathbb{N}^{|\mathcal{X}|}$ 
  \STATE \quad {\bf For} $i = 1 \dots |\mathcal{X}| - 1$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{|\mathcal{X}|}{\epsilon}}$
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_{|\mathcal{X}|}} = \alpha_{|\mathcal{X}|} + \lfloor n - \sum_{i = 1}^{|\mathcal{X}|-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \STATE {\bf return} $\tilde{\valpha}$
  \end{algorithmic}
\end{algorithm}
\begin{lem}
  Algorithm \ref{lapmech} is $\epsilon$-d.p.
  \begin{proof} By noticing that $\mathcal{X}-1$ is an upper bound to the value denoted by (\ref{eqn:eq1}). \end{proof}
\end{lem}

\subsection{Improving the sensitivity ($\ilapmech$)}
\label{sec_ell1_hist}
By seeing the parameters of the distributions as histograms, we can give a tighter upper bound to the value denoted by (\ref{eqn:eq1}).
Indeed a constant will do. Algorithm $\ilapmech$ in Algorithm \ref{ilapmech} takes advantage of this fact. 
Indeed, when $\mathcal{X}=2$, 1 is an upper bound for the value of (\ref{eqn:eq1}), while when $\mathcal{X}>2$, 2 is.
In Algorithm \ref{ilapmech} the algorithm $\ilapmech$ is presented.
\begin{algorithm}
  \caption{$\ilapmech$}
  \label{ilapmech}
  \begin{algorithmic}
  \INPUT $\dataobs\in\mathcal{X}^n$, $\dirichlet{\valpha}$
  \STATE \quad {\bf let} $\dirichlet{\valpha'} = \inferd{\valpha}{\vtheta}{\dataobs}$
  \STATE \quad ${\bf let}\ k=\left \{
        \begin{array}{lll}          
          1&\text{if}& |\mathcal{X}|=2\\
          2&\text{otherwise}
        \end{array}
      \right .\ $
  \STATE \quad {\bf Initialize} a vector $\tilde{\valpha} = (0, \dots, 0 )\in \mathbb{N}^{|\mathcal{X}|}$ 
  \STATE \quad {\bf For} $i = 1 \dots |\mathcal{X}|-1$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{k}{\epsilon}}$
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_{|\mathcal{X}|}} = \alpha_{|\mathcal{X}|} + \lfloor n - \sum_{i = 1}^{|\mathcal{X}|-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \STATE {\bf return} $\tilde{\valpha}$
  \end{algorithmic}
\end{algorithm}
\begin{lem}
  Algorithm \ref{ilapmech} is $\epsilon$-d.p.
  \begin{proof} By noticing that 1 and 2 are upper bounds to the value denoted by (\ref{eqn:eq1}), respectively when $|\mathcal{X}|=2$ and $|\mathcal{X}|>2$.
  \end{proof}
\end{lem}

\subsection{A new family of mechanisms}
\label{sec_exp}
In this section, we explore  mechanism which we denote by $\expmech$. $\expmech$ is an instance
of the exponential mechanism \cite{talwar} where the scoring function is based on a metric over
probability distributions.
The idea is that $\expmech$ will sample and output a distribution with higher probability when
it is closer to the real posterior; importantly,  \emph{closer} is measured using a metric over probability
distributions as opposed to a metric over the parameters that represent it. In this work we will
use the Hellinger distance, but this approach could be extended to other $f$-divergences. We will
start by describing how the mechanism works in the case of the Beta-Binomial model noticing that
the extension to the Dirichelet-Multinomial model is straightforward.
Given a prior distribution $\betad(\alpha, \beta)$ and an observed data set: $\dataobs\in\{0,1\}^n$,
we define the follwing set  where the mechanism will sample from:
\[
  \candidateset=\{\bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta),\dataobs)\mid \dataobs \in \{0,1\}^n\}.
\]
Where $\bysinfer(\binomial{n}{\theta}, \bprior,\dataobs)$ denotes the posterior beta distribution inferred starting from the parameters in input.
Notice that the above set reduces to:
\[
  \{\betad(\alpha',\beta')\mid \alpha'=\alpha + \Delta\alpha, \beta'=\beta + n-\Delta\alpha\}
\]
We use $\bysinfer(\dataobs)$ to denote the Bayesian inference process on the observed data $\dataobs$
when the other two parameters, that is prior and likelihood, are fixed.

\subsubsection{A first approach: the $\expmech$ mechanism}
\label{sec_ehd}
$\expmech(x,u,\candidateset)$  outputs an element $r\in\candidateset$ with  probability: % proportional to $\exp(\frac{\epsilon u(x,r)}{2 GS})$.
\[
%\underset{z \thicksim \expmech(x,u,\candidateset) }{\Pr}[z=r] = \frac
\frac
  {\exp(\frac{\epsilon\cdot \ux{r}}{2 GS})}
{\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot \ux{r'}}{2\cdot GS})},
\]
where $u(x,r)$ is the Hellinger scoring function: $-\hellinger(\bysinfer(\dataobs), r)$, and $GS$ is the global sensitivity of the scoring function, defined as:
\begin{equation*}
GS = 
\max_{\dataobs' \in \datauni^n: \\ \adj{\dataobs}{\dataobs'}}\max_{\{r\in \mathcal{R}\}}
|\hellinger(\bysinfer(\dataobs), r) - \hellinger(\bysinfer(\dataobs'), r)|
\end{equation*}
%
  \begin{algorithm}
  \caption{$\expmech$ in Beta-binomial model}
  \label{mech:expmech}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \mathcal{X}^n$, inference algorithm: $\bysinfer$, 
  prior: $\betad(\alpha, \beta)$, likelihood: $\binomial{n}{\theta}$, $\epsilon$
  \STATE \quad {\bf let} $\betad(\alpha', \beta') = \bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$. $GS$ be the global sensitivity of scoring function.  
  \STATE \quad {\bf Initialize} a vector $Pr = (0, \dots, 0 )$ of $|\candidateset|$ length.
  \STATE \quad {\bf For} $r$ in $\candidateset$:
  \STATE \quad \quad calculate the outputting probability of $r$.
  \STATE \quad \quad $Pr[r] = 
  {\exp(\frac{\epsilon\cdot \ux{r}}{2 GS})}
/ {\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot u(\dataobs, r')}{2\cdot GS})}$
\STATE \quad sample $z$ from $\candidateset$ by probability distribution $Pr$.
  \STATE {\bf return} $z$
  \end{algorithmic}
  \end{algorithm}

Lemma \ref{??} gives us a finite bound on $GS$, while lemma \ref{lem_exp_private} proves privacy $\epsilon$ differential privacy of $\expmech$.

\subsubsection{A second approach based on local sensitivity: $\lexpmech$}
\label{sec_ehdl}
Figure \ref{fig_theory} show that  the accuracy of $\expmech$ is poor if compared with the one of $\ilapmech$. This is due to the fact that
$GS$ is relatively big and hence the variance of the distribution $\expmech$ ends up also being too big. In order to solve this issue we
will proceed in two steps. We will first define $\lexpmech$, which instead of using the global sensitivity, will use the local sensitivity to calibrate
the noise. This is known to be a non private approach, but it will guide us towards new insights.
We will then proceed to use a smooth bound the local sensitivity to
calibrate the noise.

  \begin{algorithm}
  \caption{$\lexpmech$ in Beta-binomial model}
  \label{mech:expmech}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \mathcal{X}^n$, inference algorithm: $\bysinfer$, 
  prior: $\betad(\alpha, \beta)$, likelihood: $\binomial{n}{\theta}$, $\epsilon$
  \STATE \quad {\bf let} $\betad(\alpha', \beta') = \bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$. $LS$ be the local sensitivity of scoring function.  
  \STATE \quad {\bf Initialize} a vector $Pr = (0, \dots, 0 )$ of $|\candidateset|$ length.
  \STATE \quad {\bf For} $r$ in $\candidateset$:
  \STATE \quad \quad calculate the outputting probability of $r$.
  \STATE \quad \quad $Pr[r] = 
  {\exp(\frac{\epsilon\cdot \ux{r}}{2 LS})}
/ {\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot \ux{r'}}{2\cdot LS})}$
\STATE \quad sample $z$ from $\candidateset$ by probability distribution $Pr$.
  \STATE {\bf return} $z$
  \end{algorithmic}
  \end{algorithm}

$\lexpmech$ shares candidate set and utility function with $\expmech$,
and outputs a result $r \in \mathcal{R}$ with probability:
\[
%\underset{z \thicksim \lexpmech(x,u,\candidateset)}{\Pr}[z=r] = \frac
\frac
  {\exp(\frac{\epsilon \cdot \ux{r}}{2\cdot LS(\dataobs)})}
{\sum_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot \ux{r'}}{2\cdot LS(\dataobs)})},
\]
where $LS(\dataobs)$ is the local sensitivity of the scoring function, defined as:
\begin{equation}\label{eq:local_sensitivity_scoring}
LS(\dataobs)=\max_{\dataobs' \in \datauni^n: \\ \adj{\dataobs}{\dataobs'}, r\in \mathcal{R}}\lvert \hellinger(\bysinfer(\dataobs'), r) - \hellinger(\bysinfer(\dataobs'), r)\rvert 
\end{equation}
The above can be proven to simplify to:
\begin{lem}
\label{lem_loc_opt}
\[
LS(\dataobs) = \max\limits_{\dataobs' \in \datauni^n:\adj{\dataobs}{\dataobs'}} \hellinger(\bysinfer(\dataobs), \bysinfer(\dataobs')).
\]
\end{lem}
The interested reader can check the proof in the Appendix.
